Pteropod biomarker analysis

### 2018-10-09 doc/reviewer2_response2.Rmd

We appreciate the concerns of the reviewer and we hope that our response here provides sufficient clarity on the intent of our analyses.  First and most importantly, we reiterate that the analysis in Figure 5 was presented to demonstrate exactly the issue that the reviewer notes and that the results jusified our conservative approach to screen collinear variables in the following analysis (i.e., those for Figs. 6 and 7).  Therefore, we feel there is value in presenting the correlations among all variables to emphasize the complexity of the system and to provide readers with an overview of associations that were more rigorously evaluated in the follow-up analysis.  The key distinction between figure 5 and the follow-up analyses is that we never claimed the results in figure 5 were causative.  Our conservative statistical design in the follow-up analyses provided us with a more solid foundation to infer mechanistic relationships.

We also emphasize that the text noted by the reviewer on p. 6, line 240 was not the analysis used for Figure 5.  The pairwise comparisons using linear models were the follow-up analyses that exluded collinear variables.  The stars in figure 5 are simply the significance of correlation coefficients (i.e., an association).

We feel there is still some disagreement about presenting correlation results in Figure 5 among multiple variables that co-vary, such as the parameters pH, pCO2, omega, and O2.  We recreated Figure 5 using the first principal component axis of a PCA of all environmental variables.  Nearly 100% of the variance was described by the first three axes: axis one explained the OA parameters pCO2, pH, O2, and Omega, axis 2 explained temperature and latitude, and axis 3 explained fluorescence.

![](prcompex_files/figure-html/pcaplo.png)
Figure: Results of principal components analysis of the environmental data.  The first plot shows the estimated axis scores for the first and second principal components and the second plot shows the scores for the second and third principal components.  The first three axes explained nearly 99% of the variation among the observations.  Length and direction of each vector shows the relative loadings of each variable with the axes.

The first axis above was used in place of the four correlated OA parameters in the original Figure 5:

![](prcompex_files/figure-html/corploPC1.png)
Figure: Modified correlation matrix of environmental variables, cellular response endpoints, and population (abundance) and physiological response endpoints for pteropods.  Correlated environmental variables related to ocean acidification were combined into a single principal component axis represented by PC1.  Darker red values are strong positive correlations and darker purple values are strong negative correlations.  Significance values at alpha = 0.05 are shown by stars (p < 0.05 \*, p < 0.005 \*\*).

### 2019-01-21 biomark.Rmd
added oracvlpx, len, o2 v ara for abund
replaced error message in app with validate message

@@

  if(is.null(trm())) stop("Model is too simple, you broke me Nina!")

@@

  # observed data
  toplo <- md()$model
  if(ncol(toplo) < 3) stop("Model is too simple, you broke me Nina!")

>>>>

## ORACvLPX
```{r}
knitr::kable(filter(mods, pte_lab %in% 'ORACvLPX'), caption = 'ORACvLPX')
```

@@

  validate(
    need(!is.null(trm()), 'Cannot show plots for models with one predictor')
  )

@@

  # observed data
  toplo <- md()$model
  req(ncol(toplo) >= 3)
